Stability of variational quantum dynamics with Gaussian approximants

### 2013-07-24 paper.tex
writing

Prior to Glauber, Bargmann pointed out that the Hilbert space of wavefunctions is isomorphic to the space of entire functions of a complex variable, under the inner product
$$\langle g,f\rangle=\int g*(z)f(z)\,d²z.$$
He introduced the isometry
$$\langle Γ^+|ψ\rangle(α)=π^{-½}\langle α|ψ\rangle,$$
which Glauber showed was a pseudoinverse of $|Γ\rangle$, in the sense that
$$|Γ\rangle\langle Γ^+|=1.$$
The other combination, $\langle Γ^+|Γ\rangle$, is an orthogonal projector from the functions of a complex variable onto the entire functions, under the Bargmann inner product.  It follows that $f=\langle Γ^+|ψ\rangle$ is the function with least norm that satisfies $|Γ\rangle f=|ψ\rangle$.  As shown above, $\|f\|=\| |ψ\rangle\|$.

@@

  Our expansion function is 
$$g(α)=f(α)e^{-½|α|²}.$$
If $f$ has the Taylor expansion
$$f(z)=∑_n c_n{zⁿ\over(n!)^½},$$
then 
$$Gg = ∑_n c_n|n〉.$$
The functions $zⁿ$ for $n≥0$ span the entire functions (or their span is a dense subset, or something like that), so let's find the norm of 
$$g(α)=e^{-½|α|²}{αⁿ\over(n!)^½},$$
the Glauber preimage of $|n〉$.  We have 
$$|g(α)|²={1\over n!}e^{-|α|²}|α|^{2n},$$
so the squared norm is
$$\|g\|²={1\over n!}\int ₀^∞2πr^{2n+1}e^{-r²}\,dr={2π\over n!}\int ₀^∞r^{2n+1}e^{-r²}\,dr.$$
According to Wolfram,
$$\int ₀^∞r^{2n+1}e^{-r²}\,dr={n!\over 2},$$
so
$$\|g\|=π^½.$$

@@

What is the appropriate norm to use?  In the case that $χ$ is a characteristic function on the domain $D$, it is presumably the truncated Bargmann norm
$$〈g,f〉=\int_D e^{-|α|²}g*(α)f(α)\,d²α=\int e^{-|α|²}χ(α)g*(α)f(α)\,d²α.$$
Note that this is only an inner product when $g$ and $f$ are restricted to $D$.  In this case, we wish to minimise 
$$\|f_n\|²=2π\int_0^b re^{-r²}f²_n(r)\,dr,$$
given 
$$2\sqrt π(n!)^{-½}\int_0^b r^{n+1}e^{-r²}f_n(r)dr=1.$$
If one considers changing $f$ by $Δa$ in an interval around $a$, the Lagrange condition is
$$re^{-r²}f_n(r)\propto r^{n+1}e^{-r²},$$
or 
$$f_n(r)\propto rⁿ.$$
The analytic functions still win.  
In this case, the integrals are identical, so, for the preimage of $|n〉$,
$$\|f_n\|²=\sqrt{πn!},$$
and, presuming $|n〉$ is the singular vector, the singular value is
$$σ_n=(πn!)^{-1/4}.$$
I've done something wrong, because the result doesn't depend on $b$.

@@

\centerline{\XeTeXpicfile svgrid.pdf width \hsize}

@@

\centerline{\XeTeXpicfile rsv2.pdf width 0.5\hsize
\XeTeXpicfile rsv3.pdf width 0.5\hsize}
\centerline{\XeTeXpicfile rsv4.pdf width 0.5\hsize}

@@

@@

The norm of $Gg$ is 1, so this suggests the singular values of $G$ are $π^{-½}$.  This can also be seen from the expansions of coherent states.  The normalised entire expansion of $|α〉$ is
$$g(β*)=〈β|α〉=e^{-½|α|²-½|β|²+β*α},$$
whence
$$\|g\|²=\int |g(β)|²\,d²β=\int e^{-|β-α|²}\,d²β=π.$$
The norm of $|α〉$ is of course 1.  If we used the Bargmann space, we could easily absorb the $π$ into the inner product, and make $G$ a generalised projector.

>>>>

Prior to Glauber, Bargmann pointed out that the Hilbert space of wavefunctions is isomorphic to the space of entire functions of a complex variable, under the inner product
$$\langle g,f\rangle=\int g*(z)f(z)\,d²z.$$
He introduced the isometry
$$\langle Γ^+|ψ\rangle(α)=π^{-½}\langle α|ψ\rangle,$$
which Glauber showed was a pseudoinverse of $|Γ\rangle$, in the sense that
$$|Γ\rangle\langle Γ^+|=1.$$
The other combination, $\langle Γ^+|Γ\rangle$, is an orthogonal projector from the functions of a complex variable onto the entire functions, under the Bargmann inner product.  It follows that $f=\langle Γ^+|ψ\rangle$ is the function with least norm that satisfies $|Γ\rangle f=|ψ\rangle$.  As shown above, $‖f‖=‖ |ψ\rangle‖$.

@@

  Our expansion function is 
$$g(α)=f(α)e^{-½|α|²}.$$
If $f$ has the Taylor expansion
$$f(z)=∑_n c_n{zⁿ\over(n!)^½},$$
then 
$$Gg = ∑_n c_n|n〉.$$
The functions $zⁿ$ for $n≥0$ span the entire functions (or their span is a dense subset, or something like that), so let's find the norm of 
$$g(α)=e^{-½|α|²}{αⁿ\over(n!)^½},$$
the Glauber preimage of $|n〉$.  We have 
$$|g(α)|²={1\over n!}e^{-|α|²}|α|^{2n},$$
so the squared norm is
$$‖g‖²={1\over n!}\int ₀^∞2πr^{2n+1}e^{-r²}\,dr={2π\over n!}\int ₀^∞r^{2n+1}e^{-r²}\,dr.$$
According to Wolfram,
$$\int ₀^∞r^{2n+1}e^{-r²}\,dr={n!\over 2},$$
so
$$‖g‖=π^½.$$

@@

What is the appropriate norm to use?  In the case that $χ$ is a characteristic function on the domain $D$, it is presumably the truncated Bargmann norm
$$〈g,f〉=\int_D e^{-|α|²}g*(α)f(α)\,d²α=\int e^{-|α|²}χ(α)g*(α)f(α)\,d²α.$$
Note that this is only an inner product when $g$ and $f$ are restricted to $D$.  In this case, we wish to minimise 
$$‖f_n‖²=2π\int_0^b re^{-r²}f²_n(r)\,dr,$$
given 
$$2\sqrt π(n!)^{-½}\int_0^b r^{n+1}e^{-r²}f_n(r)dr=1.$$
If one considers changing $f$ by $Δa$ in an interval around $a$, the Lagrange condition is
$$re^{-r²}f_n(r)\propto r^{n+1}e^{-r²},$$
or 
$$f_n(r)\propto rⁿ.$$
The analytic functions still win.  Let $f_n(r)=rⁿ$.  Then
$$‖f_n‖²=2π\int_0^b r^{2n+1}e^{-r²}\,dr$$
and
$$‖ |Γ_χ〉f_n‖=2\sqrt π(n!)^{-½}\int_0^b r^{2n+1}e^{-r²}dr,$$
so we have
$$σ_n={‖ |Γ_χ〉f_n‖\over ‖f_n‖}=\sqrt{2\over n!}
	\left(\int_0^b r^{2n+1}e^{-r²}dr\right)^½.$$
According to Wolfram,
$$\int_0^b r^{2n+1}e^{-r²}dr=½\left(Γ(n+1)-Γ(n+1,b²)\right),$$
where $Γ$ is Mathematica's version of the incomplete gamma function.  Therefore, recalling that $n!=Γ(n+1)$,
$$σ_n=\left(1-{Γ(n+1,b²)\over Γ(n+1)}\right)^½.$$
Unfortunately, Matlab defines the incomplete gamma function in a different way than Mathematica does.  Fortunately, that Matlab function, {\tt gammainc}, is exactly the square of the right-hand side of the above equation.

This all seems wrong, because the incomplete gamma function varies the wrong way with $n$.  It becomes small as $n→0$.

@@

\topinsert \XeTeXpicfile svgrid.pdf width \hsize \endinsert

@@

\topinsert \XeTeXpicfile rsv2.pdf width 0.5\hsize
\XeTeXpicfile rsv3.pdf width 0.5\hsize \endinsert
\topinsert \XeTeXpicfile rsv4.pdf width 0.5\hsize \endinsert

@@

%% Sampling on random grids

\topinsert \XeTeXpicfile gweight.pdf width \hsize 
The weights $w_i$ required in order that the integrals of the functions $u_i(x)=e^{-(x-x_i)²}$ are given by the sums $∑_j w_j u_i(x_j)$.  The points $x_i$ are normally distributed (top), or evenly spaced (bottom).  Note how nearby points tend to cancel each others' samples.  I expect experts in Chebyshev sampling would not be surprised by this, but I am not yet such an expert.
\endinsert

@@

The norm of $Gg$ is 1, so this suggests the singular values of $G$ are $π^{-½}$.  This can also be seen from the expansions of coherent states.  The normalised entire expansion of $|α〉$ is
$$g(β*)=〈β|α〉=e^{-½|α|²-½|β|²+β*α},$$
whence
$$‖g‖²=\int |g(β)|²\,d²β=\int e^{-|β-α|²}\,d²β=π.$$
The norm of $|α〉$ is of course 1.  If we used the Bargmann space, we could easily absorb the $π$ into the inner product, and make $G$ a generalised projector.
